---
title: "MATH5472 Final Project: Demo for Replication of SMASH"
header-includes:
  - \usepackage[ruled,vlined,linesnumbered]{algorithm2e}
output:
  html_document:
    df_print: paged
  pdf_document: default
date: '2023-12-15'
author: Ruizhao Huang
---

### $\textbf{Load Packges and Source Codes}$

```{r, fig.width=10, fig.height=8}
rm(list=ls())
setwd("/Users/ruizhaohuang/Desktop/hws/MATH5472_Final_Project/Codes")
suppressPackageStartupMessages(library(wavethresh))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(smashr))
suppressPackageStartupMessages(library(ashr))
source('./src/signals.R')
source('./src/smash_em.R')
```

\

### $\textbf{Generate Normal Data Using Spike Mean Function}$

```{r, fig.width=10, fig.height=8}
set.seed(1)
f_num <- 4
N <- 2^f_num
t <- (1:N)/N
mu <- spike.fn(t,"mean")
sigma.ini <- sqrt(cblocks.fn(t,"var"))
sd.fn <- sigma.ini/mean(sigma.ini) * sd(mu)/3
z <- rnorm(N, mu, sd.fn)
```

\

### $\textbf{Apply Non-Decimated Wavelet Transform}$
```{r, fig.width=10, fig.height=8}
I_N <- diag(1, N)
W_all <- NULL
for (i in 1:N){
    w_tmp <- wavethresh::wd(diag(1, N)[i, ], type="station", filter.number=10, family="DaubLeAsymm")$D
    W_all <- cbind(W_all, w_tmp)
}
w_all <- as.numeric(W_all %*% z)
idx <- seq(1, dim(W_all)[1]+1, N)
W <- list()
w <- list()
for (i in 1:f_num){
    W[[i]] <- W_all[idx[i]:(idx[i+1]-1), ]
    w[[i]] <- w_all[idx[i]:(idx[i+1]-1)]
}
```

\

### $\textbf{Replication of SMASH Using My Own Implementation of ASH}$
Note that it may took a while to run, because the slow convergence of standard EM algorithm.
```{r, fig.width=10, fig.height=8}
res1 = smash_em(z, w, W, s)
```

\

### $\textbf{Replication of SMASH using ash Function in ashr Package}$

Note that it may took a while to run, because the slow convergence of standard EM algorithm, especially when number of observation is large.

```{r, fig.width=10, fig.height=8}
# initial estimate of s
lag_z <- lag(z, 1) %>% replace(1, tail(z, 1))
lead_z <- lead(z, 1) %>% replace(N, head(z, 1))
s = sqrt(0.5 * ((z - lag_z)^2 + (z - lead_z)^2))

iter <- 2
mu_est_mat <- matrix(0, nrow=(iter+1), ncol=N)
s_est_mat <- matrix(0, nrow=(iter+1), ncol=N)
s_est_mat[1, ] <- s
for (k in 2:(iter+1)){
  # estimate mu
  trans_mu <- data_trans_mu(z, s_est_mat[k-1, ], W)
  est_hat <- matrix(0, nrow=f_num, ncol=N)
  for(i in 1:f_num){
    ash_res <- ashr::ash(betahat=w[[i]], sebetahat=trans_mu$tilde_s[[i]],
                         mixcompdist='normal', optmethod='mixEM')
    est_hat[i, ] <- as.numeric(t(W[[i]]) %*% get_pm(ash_res))
  }
  mu_est_mat[k, ] <- colMeans(est_hat)
   
  # estimate s
  trans_s <- data_trans_var(z, mu_est_mat[k, ], W)
  est_hat <- matrix(0, nrow=f_num, ncol=N)
  for(i in 1:f_num){
    ash_res <- ashr::ash(betahat=trans_s$tilde_w[[i]],
                         sebetahat=trans_s$tilde_psi[[i]], 
                         mixcompdist='normal', optmethod='mixEM')
    est_hat[i, ] <- as.numeric(t(W[[i]]) %*% get_pm(ash_res))
  }
  s_est_mat[k, ] <- colMeans(est_hat)
  s_est_mat[s_est_mat <= 0] <- 1e-4
  s_est_mat[k, ] <- sqrt(s_est_mat[k, ])
}

res2 = mu_est_mat[(iter+1), ]
```

\

### $\textbf{Directly call smash Function in smashr Package}$
Note that it may took a while to run, because the slow convergence of standard EM algorithm.
```{r, fig.width=10, fig.height=8}
res3 <- smash(z, family="DaubLeAsymm", filter.number = 10)
```

\

### $\textbf{Comparison of My Replication with Existing Package}$
```{r, fig.width=5, fig.height=5}
# pdf(file = '../Report/plot/compare_smash.pdf', width = 5, height = 5)
MSE_res1 <- sum((res1$mu - mu)^2) / N
MSE_res2 <- sum((res2 - mu)^2) / N
MSE_res3 <- sum((res3 - mu)^2) / N
MSE <- tibble(Algo = c('My_Rep', 'My_Rep+ash', 'Package'), 
              MSE = c(MSE_res1, MSE_res2, MSE_res3))

ggplot(data=MSE, aes(x=Algo, y=MSE, fill=Algo)) +
  geom_bar(stat="identity", width=0.6) +
  theme_minimal()
# dev.off()
```

